---
title: "R Notebook"
output: html_notebook
---
```{r}
library(readr)
library(readxl)
library(tidyr)
library(dplyr)
library(immunarch)
library(data.table)
```

Load subsampled data
```{r}
# this object was created in the immunarch_analysis script
sub <- readRDS(file="sub.Rdata")
```

# Create dataframe and perform feature filtering

generate dataframe: features x observations. 
proportions of public clonotype sharing
```{r}
public_prop <- pubRep(sub, .quant = "prop", .verbose = F, .col = "aa")
```

```{r}
pub_prop_df = as.data.frame(public_prop)
pub_prop_df[is.na(pub_prop_df)] = 0
pub_prop_df_wos <- subset(pub_prop_df, select = -c(Samples) )
```

transpose dataframe: observations x features
```{r}
merged <- pub_prop_df_wos
merged_t <- setNames(data.frame(t(merged[,-1])), merged[,1])
```

turn data into matrix
```{r}
data_m <- data.matrix(merged_t)
```

# Method 1: 600 most shared features
select top 600 features
1. Count how many non zero cells in each column. (How many times does each feature appear?)
```{r}
samples_count <- colSums(data_m != 0)
```

2. indices of the sorted vector from high to low and keep top 600.
```{r}
feature_selection <- order(samples_count, decreasing = T)[1:600]
```

3. select those indices from data table.
```{r}
data_fs <- merged_t[,feature_selection]
```

export to csv
```{r}
write.csv(data_fs,"./data_600_ab.csv", row.names = T)
```


# Method 2: top features from databases

Database: VDJdb
```{r}
file <- "./SearchTable.tsv"
vdjdb <- read_tsv(file)

vdj_2 <- vdjdb[,-c(1,13,14,15,16)]

vdj_uniq <- unique(vdj_2)
```

```{r}
top_uniq <- sort(table(vdj_uniq$CDR3), decreasing = T)[1:10]
top_uniq
```

Database: McPAS
```{r}
path <- "./mc.10.9.22.xlsx"
mcpas <- read_excel(path)
```

```{r}
cdr3_mcpas <- sort(c(table(mcpas$CDR3.alpha.aa), table(mcpas$CDR3.beta.aa)), decreasing = T)[-c(1,2)]
top_mcpas <- cdr3_mcpas[1:10]
top_mcpas
```

Database: Trust4 TCGA
```{r}
tcga_file <- "./trust4_tcga_deidentified_report.tsv"
tcga_trust4 <- read_tsv(tcga_file, col_names = c("sample", "chain", "cdr3", "subchain", "count"))

# subset tcr data
trust_tcr <- subset(tcga_trust4, chain %in% c("TRA", "TRB"))
```

```{r}
top_tcga <- sort(table(trust_tcr$cdr3), decreasing = T)[2:11]
top_tcga
```

```{r}
# overlapping TCRs from all 3 sets
Reduce(intersect, list(names(top_uniq), names(top_mcpas), names(top_tcga)))
```

concatenate all tcrs from vdjdb, mcpas and tcga
```{r}
tcrs <- c(names(top_uniq), # vdjdb top 10
          names(top_mcpas), # mcpas top 10 
          names(top_tcga)) # tcga top 10
```

```{r}
# from 30 down to 25 bc of overlap between databases.
uniq_tcrs <- unique(tcrs)
uniq_tcrs
```

```{r}
# check which of the 25 features are present in our dataset
colnames <- colnames(merged_t)
intersection <- intersect(uniq_tcrs, colnames) # only 16 of 25!
```

select features from data table.
```{r}
data_fs <- merged_t[,intersection]
```

```{r}
write.csv(data_fs,"./data_16_ab.csv", row.names = T)
```

