---
title: "R Notebook"
output: html_notebook
---

```{r warning = FALSE}
library("immunarch")
library("purrr")
library("stringr")
library(forcats)
library(data.table)
library(ggpubr)
library(scales)
library(plotly)
```

All Samples (H+OC) TRA and TRB 
```{r}
#### BLOOD
immdata<-repLoad("/path/to/processed/data/")
```

Low clones
```{r}
clones<-repExplore(immdata$data, .method = "clones")
blood_order = arrange(clones, Clones)
rownames(blood_order) = c()
head(blood_order)
```

# Subsample
1. Omit 21_H, 49_H
```{r}
immdata$data[which(names(immdata$data) %in% c("21_A_B_H", "49_A_B_H"))] <- NULL
```

2. subsample
```{r}
sub = repSample(immdata$data, .method = "downsample", .n = 24840)
```

```{r}
# this object will be used in the feature_filtering script as well.
saveRDS(sub, "sub.Rdata")
```

# Initial data analysis
```{r}
sub_h <- sub[grepl("H",names(sub))]
sub_oc <- sub[grepl("OC",names(sub))]
```

analysis after subsampling
sub_h and sub_oc contain both TRA and TRB

0. unique
1. top
2. rare
3. diversity
    a. gini
    b. True diversity "div"
    c. Inverse Simpson index "inv.simp"
    d. Gini-Simpson index "gini.simp"
4. overlap


0. unique
```{r}
uniq_h<-repExplore(sub_h, .method = "volume")

uniq_df_h <- data.frame(uniq_h)

plt_uniq_h <- ggplot(uniq_df_h,aes(x=Sample,y=Volume))+
  geom_bar(stat='identity',position='stack')+
  labs(title = "Unique clonotypes in Healthy Donors peripheral blood", x = "Sample", y = "Count") + 
  #theme(axis.text.x = element_text(angle = 90))+
  theme_bw()+
  theme(axis.text.x = element_blank(), panel.grid = element_blank())+
  scale_y_continuous(labels = comma)

plt_uniq_h
```

```{r}
uniq_oc<-repExplore(sub_oc, .method = "volume")
vis(uniq_oc)
uniq_df_oc <- data.frame(uniq_oc)

plt_uniq_oc <- ggplot(uniq_df_oc,aes(x=Sample,y=Volume))+
  geom_bar(stat='identity')+
  labs(title = "Unique clonotypes in OC peripheral Blood", x = "Sample", y = "Count") +
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90))+
  theme(axis.text.x = element_blank(),  panel.grid = element_blank())+
  scale_y_continuous(labels = comma)

plt_uniq_oc
```

t test
```{r}
uniq_y_h <- uniq_h$Volume
uniq_y_oc <- uniq_oc$Volume

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(uniq_y_h, uniq_y_oc)$p.value # 0.73
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(uniq_y_h, uniq_y_oc, var.equal = var.equal)

# p value = 0.19. no significant difference.
```


1. top
```{r}
top_h <- repClonality(sub_h, .method = "top", .head= c(10,100,1000,3000,10000,30000)) 

top_df <- data.frame(top_h)
top_n <- mutate(top_df, "[11:100)" = X100 - X10, 
                  "[101:1000)" = X1000 - X100, 
                  "[1001:3000)" = X3000 - X1000, 
                  "[3001:10000)" = X10000 - X3000,
                  "[10001:30000)" = X30000 - X10000,
                  "Patient_ID" = rownames(top_df),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))

top_n1 <- top_n[,c("Patient_ID","X10","[11:100)", "[101:1000)","[1001:3000)","[3001:10000)","[10001:30000)")] 
top_n1 = top_n1 %>% rename( "[1:10)" = "X10")

test <- reshape2::melt(top_n1)
plt <- ggplot(test,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Top clonal proportion HD", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype indices"))+
  scale_fill_brewer(palette = "RdYlBu")

plt
```

```{r}
top_oc <- repClonality(sub_oc, .method = "top", .head= c(10,100,1000,3000,10000,30000)) 

top_df_oc <- data.frame(top_oc)
top_n_oc <- mutate(top_df_oc, "[11:100)" = X100 - X10, 
                  "[101:1000)" = X1000 - X100, 
                  "[1001:3000)" = X3000 - X1000, 
                  "[3001:10000)" = X10000 - X3000,
                  "[10001:30000)" = X30000 - X10000,
                  "Patient_ID" = rownames(top_df_oc),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))

top_n1_oc <- top_n_oc[,c("Patient_ID","X10","[11:100)", "[101:1000)","[1001:3000)","[3001:10000)","[10001:30000)")] 
top_n1_oc = top_n1_oc %>% rename( "[1:10)" = "X10")

test_oc <- reshape2::melt(top_n1_oc)
plt_oc <- ggplot(test_oc,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Top clonal proportion OC", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype indices"))+
  scale_fill_brewer(palette = "RdYlBu")

plt_oc
```

t test
```{r}
top_y_h <- top_n1$`[1:10)`

top_y_oc <- top_n1_oc$`[1:10)`

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(top_y_h, top_y_oc)$p.value # 0.74
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(top_y_h, top_y_oc, var.equal = var.equal)

# p-value = 0.057 no significant difference
```


2. rare
```{r}
rare_h <- repClonality(sub_h, .method = "rare")

rare_df_h <- data.frame(rare_h)
rare_h_n <- mutate(rare_df_h, "2 - 3" = X3 - X1, 
                  "4 - 10" = X10 - X3, 
                  "11 - 30" = X30 - X10, 
                  "31 - 100" = X100 - X30,
                  "101 - MAX" = MAX - X100,
                  "Patient_ID" = rownames(rare_df_h),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))
rare_h_n1 <- rare_h_n[,c("Patient_ID","X1","2 - 3", "4 - 10", "11 - 30","31 - 100","101 - MAX")]
rare_h_n1 = rare_h_n1 %>% rename( "1" = "X1")
test_rare_h <- reshape2::melt(rare_h_n1)

plt_rare_h <- ggplot(test_rare_h,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Rare clonal proportion HD", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype counts"))+
  scale_fill_brewer(palette = "RdYlBu")

plt_rare_h
```


```{r}
rare_oc <- repClonality(sub_oc, .method = "rare")

rare_df_oc <- data.frame(rare_oc)
rare_n_oc <- mutate(rare_df_oc, "2 - 3" = X3 - X1, 
                  "4 - 10" = X10 - X3, 
                  "11 - 30" = X30 - X10, 
                  "31 - 100" = X100 - X30,
                  "101 - MAX" = MAX - X100,
                  "Patient_ID" = rownames(rare_df_oc),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))
rare_n1_oc <- rare_n_oc[,c("Patient_ID","X1","2 - 3", "4 - 10", "11 - 30","31 - 100","101 - MAX")]
rare_n1_oc = rare_n1_oc %>% rename( "1" = "X1")

test_rare_oc <- reshape2::melt(rare_n1_oc)

plt_rare_oc <- ggplot(test_rare_oc,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  theme(axis.text.x = element_blank(), legend.position = "none", complete = F, panel.grid = element_blank())+
  labs(title = "Rare clonal proportion OC", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype counts"))+
  scale_fill_brewer(palette = "RdYlBu")

plt_rare_oc
```

t test
```{r}
rare_y_h = rare_h_n1[,2]

rare_y_oc = rare_n1_oc[,2]

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(rare_y_h, rare_y_oc)$p.value # 0.62 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(rare_y_h, rare_y_oc, var.equal = var.equal)

# p-value = 0.57, no significant difference
```


3a. diversity- gini
```{r}
gini_h <- repDiversity(sub_h, .method = "gini", .verbose = F)

gini_df_h = data.frame(gini_h)
gini_df_h$sample <- rownames(gini_df_h)
gini_df_h$Num = as.numeric(map(str_split(gini_df_h$sample, regex("_"),2),1))
gini_df_h$Patient_ID = fct_reorder(gini_df_h$sample, gini_df_h$Num)

plt_gini_h<- ggplot(gini_df_h, aes(Patient_ID, gini_h)) + #, fill = sample
  geom_col()+
  theme_bw()+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Diversity HD", x = "Sample", y = "gini index")+
  theme(legend.position = "none")

plt_gini_h
```

```{r}
gini_oc <- repDiversity(sub_oc, .method = "gini", .verbose = F)

gini_df_oc = data.frame(gini_oc)
gini_df_oc$sample <- rownames(gini_df_oc)
gini_df_oc$Num = as.numeric(map(str_split(gini_df_oc$sample, regex("_"),2),1))
gini_df_oc$Patient_ID = fct_reorder(gini_df_oc$sample, gini_df_oc$Num)

plt_gini_oc<- ggplot(gini_df_oc, aes(Patient_ID, gini_oc)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Diversity OC", x = "Sample", y = "gini index")+
  theme(legend.position = "none")

plt_gini_oc
```

t test
```{r}
gini_y_h = gini_df_h$gini_h

gini_y_oc = gini_df_oc$gini_oc

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(gini_y_h, gini_y_oc)$p.value # 0.73 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(gini_y_h, gini_y_oc, var.equal = var.equal)

# p-value = 0.137, no significant difference
```


3b. diversity - True diversity "div"
```{r}
true_d_h <- repDiversity(sub_h, .method = "div", .verbose = F)

true_d_h$Num = as.numeric(map(str_split(true_d_h$Sample, regex("_"),2),1))
true_d_h$Patient_ID = fct_reorder(true_d_h$Sample, true_d_h$Num)

plt_trued_h<- ggplot(true_d_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "True Diversity HD", x = "Sample", y = "True Diversity")+
  theme(legend.position = "none")

plt_trued_h
```

```{r}
true_d_oc <- repDiversity(sub_oc, .method = "div", .verbose = F)

true_d_oc$Num = as.numeric(map(str_split(true_d_oc$Sample, regex("_"),2),1))
true_d_oc$Patient_ID = fct_reorder(true_d_oc$Sample, true_d_oc$Num)

plt_trued_oc<- ggplot(true_d_oc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "True Diversity OC", x = "Sample", y = "True Diversity")+
  theme(legend.position = "none")

plt_trued_oc
```

t test
```{r}
trued_y_h = true_d_h$Value

trued_y_oc = true_d_oc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(trued_y_h, trued_y_oc)$p.value # 0.41
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(gini_y_h, gini_y_oc, var.equal = var.equal)

# p-value = 0.137, no significant difference
```


3c. diversity - Inverse Simpson index "inv.simp"
```{r}
inv_h <- repDiversity(sub_h, .method = "inv.simp", .verbose = F)

inv_h$Num = as.numeric(map(str_split(inv_h$Sample, regex("_"),2),1))
inv_h$Patient_ID = fct_reorder(inv_h$Sample, inv_h$Num)

plt_inv_h<- ggplot(inv_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Inverse Simpson Index Diversity HD", x = "Sample", y = "Inverse Simpson index")+
  theme(legend.position = "none")

plt_inv_h
```

```{r}
inv_oc <- repDiversity(sub_oc, .method = "inv.simp", .verbose = F)

inv_oc$Num = as.numeric(map(str_split(inv_oc$Sample, regex("_"),2),1))
inv_oc$Patient_ID = fct_reorder(inv_oc$Sample, inv_oc$Num)

plt_inv_oc<- ggplot(inv_oc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Inverse Simpson Index Diversity OC", x = "Sample", y = "Inverse Simpson index")+
  theme(legend.position = "none")

plt_inv_oc
```

t test
```{r}
inv_y_h = inv_h$Value

inv_y_oc = inv_oc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(inv_y_h, inv_y_oc)$p.value # 0.69 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(inv_y_h, inv_y_oc, var.equal = var.equal)

# p-value = 0.048, significant difference!!
```


3d. diversity - Gini-Simpson index "gini.simp"
```{r}
simp_h <- repDiversity(sub_h, .method = "gini.simp", .verbose = F)

simp_h$Num = as.numeric(map(str_split(simp_h$Sample, regex("_"),2),1))
simp_h$Patient_ID = fct_reorder(simp_h$Sample, simp_h$Num)

plt_simp_h<- ggplot(simp_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Simpson Index Diversity HD", x = "Sample", y = "Gini-Simpson index")+
  theme(legend.position = "none")

plt_simp_h
```

```{r}
simp_oc <- repDiversity(sub_oc, .method = "gini.simp", .verbose = F)

simp_oc$Num = as.numeric(map(str_split(simp_oc$Sample, regex("_"),2),1))
simp_oc$Patient_ID = fct_reorder(simp_oc$Sample, simp_oc$Num)

plt_simp_oc<- ggplot(simp_oc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Simpson Index Diversity OC", x = "Sample", y = "Gini-Simpson index")+
  theme(legend.position = "none")

plt_simp_oc
```


t test
```{r}
simp_y_h = simp_h$Value

simp_y_oc = simp_oc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(simp_y_h, simp_y_oc)$p.value # 0.129 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(simp_y_h, simp_y_oc, var.equal = var.equal)

# p-value = 0.39, no significant difference
```

4. overlap
```{r}
public <- repOverlap(sub, .method = "public", .verbose = F, .col = "aa")
public_df = as.data.frame(public)

variable_names <- names(public_df)

# Order the variable names
ordered_names <- variable_names[order(gsub(".*_", "", variable_names), variable_names)]

# Order the dataframe columns
public_df_ordered <- public_df[, ordered_names]

# Order the dataframe rows
public_df_ordered2 <- public_df_ordered[ordered_names, ]

f_a <- plot_ly(x = colnames(public_df_ordered2), 
               y = rownames(public_df_ordered2), 
               z = as.matrix(public_df_ordered2), 
               type = "heatmap", 
               zauto = F, zmin = 0, zmax = 800) %>%
  layout(xaxis = list(tickangle=-90))

f_a <- f_a %>% layout(title = 'Repertoire Overlap Blood OC and HD',
         xaxis = list(title = 'Sample'),
         yaxis = list(title = 'Sample'))

f_a
```



